---
title: "p8105_hw1_bl3175"   
output: github_document                  
---

```{r setup, include=FALSE}
# Load required packages and set knitr chunk options
library(tidyverse)
library(moderndive)
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
data("early_january_weather")
glimpse(early_january_weather)
```

## Problem 1 

**Dataset description (with inline R)**  
This dataset has **`r nrow(early_january_weather)`** rows and **`r ncol(early_january_weather)`** columns.  
Variable names include: `r paste(names(early_january_weather), collapse = ", ")`.  
The mean temperature is **`r round(mean(early_january_weather$temp, na.rm = TRUE), 1)`** °F.  
The `time_hour` variable is of class **`r class(early_january_weather$time_hour)[1]`** (hourly timestamps).  
Humidity (`humid`) ranges from **`r round(min(early_january_weather$humid, na.rm = TRUE), 2)`** to **`r round(max(early_january_weather$humid, na.rm = TRUE), 2)`**.

## Scatterplot of temp vs time_hour (colored by humidity)

```{r fig.width=7, fig.height=4}
# Build the scatterplot and store it in object `p`
p <- ggplot(early_january_weather, aes(x = time_hour, y = temp, color = humid)) +
  geom_point(alpha = 0.6) +
  labs(
    title = "Temperature vs Time (colored by humidity)",
    x = "Time (hourly)",
    y = "Temperature (°F)",
    color = "Humidity"
  ) +
  theme_minimal()

# Print the plot
p
```

## Export the plot to the project directory

```{r}
# Create a folder to keep figures

if (!dir.exists("figs")) dir.create("figs")

# Save the plot 
ggsave("figs/temp_vs_time.png", plot = p, width = 7, height = 4, dpi = 300)
```